---
title: "Installing R and Rstudio"
author: "Michael Stevens"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Install R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
# library(imager)
# embed_png <- function(path, dpi = NULL) {
#   meta <- attr(png::readPNG(path, native = TRUE, info = TRUE), "info")
#   if (!is.null(dpi)) meta$dpi <- rep(dpi, 2)
#   knitr::asis_output(paste0(
#     "<img src='", path, "'",
#     " width=", round(meta$dim[1] / (meta$dpi[1] / 96)),
#     " height=", round(meta$dim[2] / (meta$dpi[2] / 96)),
#     " />"
#   ))
# }
# knitr::opts_chunk$set(comment = "#>", collapse = TRUE)
```


This really is the very first step to getting to grips with R. Before we can do anything, we will need to install it on your computer.  

# Installing R 

We are going to install R from the [R project website](https://www.r-project.org/). This is the main website for finding different versions of R, hearing news, reporting bugs and help with using R in general. We will be installing:

Here's the image 

 ```{r, echo = FALSE}
# embed_png("images/test.png", dpi = 10)
```

# Installing Rstudio
