---
title: "Exercise solutions"
author: "Michael Stevens"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Exercise solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## 1. Calculations

Use R to carry out the following calculations:

* 3 times 5

```{r}
3*5
```
 
* 5 to the power of 2

```{r}
5^2
```

* Add together 8 and 22 and then multiply the answer by 3

```{r}
(8 + 22)/3
```

* Subtract 2.8 from 12 and then raise the answer to the power of 1.5 plus 1.3924

```{r}
(12 - 2.8)^(1.5 + 1.3924)
```

* Divide 8 by 2.5 and then divide the answer by 3

```{r}
(8/2.5)/3
```

* Subtract 2.5 from 8 and then divide the answer by 3

```{r}
(8 - 2.5)/3
```

* Divide 2.5 by 3 and subtract the answer from 8

```{r}
8 - (2.5/3)
```

## 2. Objects

* Create an object called "x" and assign it the numerical value 73

```{r}
x <- 73
```

* Create another object called “y” which is the answer to the sum of 101 and 36

```{r}
y <- 101 + 36
```

* Multiply x and y together and assign this value to the object called z

```{r}
z <- x*y
```

* Subtract 1 from z and then calculate the 4th root (note you can raise to the ith root by raising to the power of 1/i)

```{r}
(z - 1)^(0.25)
```

## 3. Functions

Using the same objects x, y and z that were set up in the previous exercises:

* Calculate the log to the base 10 of x using the function `log10()`

```{r}
log10(10)
```

* Calculate the square root of y using the function `sqrt()`

```{r}
sqrt(y)
```

* Calculate the square root of y by raising it to the power 0.5: your answer should be the same as when you used the `sqrt()` function

```{r}
y^0.5
```

* use the function `sum()` to add x and y: you'll need to put both object names inside the brackets and separate them with a comma

```{r}
sum(x,y)
```

* use the `mean()` function to calculate the arithmetic mean (i.e. the average) of x and y. NB mean() will only work on a vector so you will need to use the `c()` function to collect x and y together.

```{r}
# collect x and y as a vector
xy_vec <- c(x,y)

mean(xy_vec)
```

* Create an object called “w” which is assigned the value 67.8953

```{r}
w <- 67.8953
```

* The function `round()` will round a number to a given number of decimal places. The default is zero so start by using this function to round w off to zero decimal places

```{r}
round(w)
```

* Now use the round function to give you the value specified by w rounded off to three decimal places

```{r}
round(w, 3)
```

* `floor()` and `ceiling()` can also be used to trim a number down to an integer: apply both of these functions to `w` and compare the output.

```{r}
# floor the value w
floor(w)

# ceiling the value w
ceiling(w)
```

## 4. Vectors

* Delete the objects x, y, z and w that you set up for the previous set of exercises using the `rm()` function

```{r}
rm(x, y, z, w)
```

* Create a vector called “x” consisting of the numbers 1,3,5 and 7

```{r}
x <- c(1, 3, 5, 7)
```

* Create a vector called “y” consisting of the numbers 2,4,6 and 8

```{r}
y <- c(2, 4, 6, 8)
```

* Subtract x from y

```{r}
y - x
```

* Create a new vector called “z” by multiplying vector x by vector y

```{r}
z <- x*y
```

* Create a new vector called “w” by taking the square root of each member of z

```{r}
w <- sqrt(z)
```

* Use the `mean()` function to calculate the arithmetic mean of the numbers in vector w

```{r}
mean(w)
```

* Use the  `median()` function to calculate the median value of the numbers in vector z

```{r}
median(z)
```

## 5. Matrices

* Create a new vector called “V1” consisting of the following numbers 1,3,5,7,9,11

```{r}
V1 <- c(1, 3, 5, 7, 9, 11)
```

* Create a matrix called “mat1” using the following function: `mat1 <- matrix(V1, nrow = 2)`

```{r}
mat1 <- matrix(V1, nrow = 2)
```

* Create a matrix called “mat2” using the same function but add an extra argument `byrow = TRUE`

```{r}
mat2 <- matrix(V1, nrow = 2, byrow = TRUE)
```

* Compare the two matrices and note how the data stored in the vector V1 are used to fill up the cells of the two matrices

```{r}
# compare mat1 and mat2
print(mat1)
print(mat2)
```

## 6. Vector indexing and logical operators

The vectors x, y, z and w refer to the vectors set up in exercise 4. 

* Find the value of the 3rd number in vector x

```{r}
x[3]
```

* Find the value of the numbers in vector y that **ARE NOT** in the 3rd position

```{r}
# use the minus symbol to show a vector without a specific entry
# give all but the third value of the vector y
y[-3]
```

* Add the 1st number in vector z to the 4th number in vector w

```{r}
z[1] + w[4]
```

* Create a new vector called “In” which consists of the numbers 1 and 4

```{r}
In <- c(1, 4)
```

* Use indexing rules and the “In” vector to create a new vector, "my_vec", which consists of the sums of the 1st and the 4th numbers of z and w

```{r}
my_vec <- c(sum(z[In]), sum(w[In]))
```

* Calculate the sum of all the numbers in y that are less than 7

```{r}
sum(y[y < 7])
```

* Calculate the mean of all the numbers in x that are greater than or equal to 3

```{r}
mean(x[x >= 3])
```

## 7. Matrix indexing

These exercises use the objects previously set up in exercise 5. 

* Multiply the second value in the first row of mat1 by the third value in the second row of mat2

```{r}
mat1[1, 2]*mat2[2, 3]
```

* Create a new vector called “V2” which consists of the numbers in the first row of mat1 added to the numbers in the second row of mat2

```{r}
V2 <- mat1[1, ] + mat2[2, ]
```

* Create a new vector called “V3” which consists of the numbers in the second column of mat2 multiplied by the mean of the numbers in the second row of mat1

```{r}
V3 <- mat2[, 2]*mean(mat1[2, ])
```

* Create a new matrix called “mat3” which consists of the first row of mat1 as the first column and then the first row of mat2 as the second column

```{r}
mat3 <- matrix(c(mat1[1, ], mat2[1, ]), 
               ncol = 2, 
               byrow = FALSE)
```

## 8. Writing functions

* Write a function that will take a number square it

```{r}
square_power <- function(x){
  value <- x^2
  return(value)
}
```

* Test it on the value 5. The test it on a vector of 15 random numbers drawn from a standard normal distribution with mean zero and standard deviation 1 (Note this can be done with the `rnorm()` function)

```{r}
# use the function on the value 5
square_power(5)

# make some random number with rnorm
random_numbers <- rnorm(n = 15, mean = 0, sd = 1)

# use these random numbers on the square function
square_power(random_numbers)
```

* Now add an argument called `power` to the function that enables the user to specify which power that the value should be raised to - i.e. instead of squaring the value, it could be raise to the power of 5 instead. Test it on your vector of random numbers again with the power set to 5 and 20.

```{r}
# rename function appropriately 
power_function <- function(x, power){
  value <- x^power
  return(value)
}

# use the function on the random numbers with the power set to 5 and 20
power_function(random_numbers, power = 5)
power_function(random_numbers, power = 20)
```

* Now add an argument called “multiply” that will let the user choose a number to multiply the value by as well. Set the default value as 2 and test it on your random numbers again this time witht he power set to 6 and the multiply set to 12 

```{r}
# rename function appropriately 
my_function <- function(x, power, multiply = 2){
  value <- multiply*x^power
  return(value)
}

my_function(random_numbers, power = 6, multiply = 12)
```


* Set up a vector called “new_vec” with a sequence of 400 numbers from 0 to 12 (using the `seq()` function) and use this and your function to plot a graph with the sequence on the x axis and the output of the function on the y axis

```{r}
new_vec <- seq(0, 12, l = 400)
y_points <- my_function(new_vec, power = 6, multiply = 12)

plot(new_vec, y_points)
```
